---
title: "Geographical Information System of NYC"
author: "Shubhada Suresh"
date: "May 4, 2018"
output: html_document
css: styles.css

---

```{r setup, include=FALSE}
library(shiny)
library(leaflet)
library(data.table)
library(geosphere)
library(tidyverse)
library(sp)
library(ggmap)
library(maptools)
library(Hmisc)
library(xlsx)
```

## Project Overview

 * NYC Subway Stations
 * NYC Restaurants
 * NYC Theaters
 * NYC Museums and Galleries
 
## Tidying data of NYC Subway stations
   Certain train stations had multiple entrance and exits hence multiple latitude and longitude.
   
   I restricted only 1 row per station as multiple location of same station will not make sense when we plot it on map.
   
   Then loaded the clean file that can be used.

```{r eval=FALSE}
SUBWAY <-
  read.csv(file = "https://raw.githubusercontent.com/shubhadasuresh/MTA-Subway/master/SUBWAY_STATION.csv", header =
             T, sep = ",")
SUBWAY$long <-
  sub('.*\\(', '', SUBWAY$the_geom)
SUBWAY$lat <- sub('.*\\s+', '', SUBWAY$long)
SUBWAY$lat <- gsub(")", "", SUBWAY$lat)
SUBWAY$long <- sub('\\s+.*$', '', SUBWAY$long)

SUBWAY$lat <- as.numeric(SUBWAY$lat)
SUBWAY$long <- as.numeric(SUBWAY$long)

write.csv(SUBWAY_cleaned,'SUBWAY.csv')

```

## Tidying data of NYC Restaurants

   * Selecting only required columns in the dataset.
   * Combining columns like DBA,BORO,BUILDING,STREET,ZIPCODE as single column Address which will be easy to feed for geocode.
   * Geocoding is the process of converting addresses (like a street address) into geographic coordinates (like latitude and longitude), which you can use to place markers on a map, or position the map
   * To use the Geocoding API, you must register your app project on the Google API Console and get a Google API key which you can add to your app or website.
   * For standard users Geocoding API supports up to 2500 requests daily. 
   * So after 12 days of continous requests, i could able to get latitude and longitude for 30000 restaurants.
   * There were so many special chracters in between the names and address which need to to be detected before feeding to geocode.
   * Many places names were little different like **1$ pizza** but in goecode it needed to be **One Dollar Pizza**.
   * Showing the short version of code here.
   
```{r eval=FALSE}
NYC_Food <- read_csv("https://nycopendata.socrata.com/api/views/xx67-kt59/rows.csv?accessType=DOWNLOAD")

NYC_Food <- NYC_Food %>% select(DBA,BORO,BUILDING,STREET,ZIPCODE,`CUISINE DESCRIPTION`)
NYC_Food <- NYC_Food[!duplicated(NYC_Food), ]


NYC_Food$x <- paste(NYC_Food$DBA,NYC_Food$BORO,NYC_Food$BUILDING,sep = ',')
NYC_Food$y <- paste(NYC_Food$STREET,"New York"," NY ",sep = ',')
NYC_Food$Address <- paste(NYC_Food$x,NYC_Food$y,NYC_Food$ZIPCODE,sep= " ")

for(i in 1:nrow(NYC_Food))
{
result <- geocode(NYC_Food$Address[i])
  NYC_Food$lon[i] <- as.numeric(result[1])
  NYC_Food$lat[i] <- as.numeric(result[2])
}

write.csv(Just,'NYC_Food.csv')

```

## Tidying data of NYC Theaters.
   * Similar to Subway station dataset.

```{r eval=FALSE}
 THEATER <- read.csv("https://raw.githubusercontent.com/shubhadasuresh/MTA-Subway/master/THEATER_cleaned.csv", header =T, sep = ",")

THEATER$long <-
  sub('.*\\(', '', THEATER$the_geom)
THEATER$lat <- sub('.*\\s+', '', THEATER$long)
THEATER$lat <- gsub(")", "", THEATER$lat)
THEATER$long <-sub('\\s+.*$','', THEATER$long)

write.csv(THEATER,'THEATER_cleaned.csv')
```

## Tidying data of NYC Museums and Galleries.

   * This dataset had all Museums and Galleries in NYC.
   * The Coordinate of latitude and longitude was combined.
   * Used Reg expression to separate it as latidude and longitude in different columns.
   * One column had website link which i used as hyperlink on map. So that on clicking we can directly see hte website of that particular Museum.
   
   
```{r eval=FALSE}
 Museums <- read.csv("https://raw.githubusercontent.com/shubhadasuresh/MTA-Subway/master/museums_galleries.csv", header =T, sep = ",")
 
 Museums$lat <- sub("\\,.*", "", Museums$Geometry.X.Y)
 Museums$long <- sub('.*,\\s*', '', Museums$Geometry.X.Y)
 
 Museums$lat <- as.numeric(Museums$lat)
 Museums$long <- as.numeric(Museums$long)
 
```

## Work in Shiny R

  * Basic intension is to provide nearby entertainment spots like Theaters, Restaurants,Museums and Galleries.
  * By now we have collected all the latitudes and longitudes of our data.
  * Now the aim is to calculate distance through whic we can determine what is near.
  * Geosphere package compute various aspects of distance, direction, area, etc.
for geographic (geodetic) coordinates.
  * Sort the coordinates in ascending order and pick 5 nearest to subrain station selected.
  * In leaflet, using different markers, Popups,Labels,colours, legend we can demonstrate **miniature version of google maps** using Shiny R.
  
  **For Demonstration Purposes** i have taken some sample of stations and Restaurants only as Shiny app hangs and times out for 30,000 rows of restaurants.
  Eg: 
  
     + Chambers St -->  W 4th St - Washington Sq (Upper)
     + Fulton Street --> Utica Ave
     + Union Sq - 14th St -->  28th St
     + Jay St - MetroTech -->  42nd St - Port Authority Bus Term
     + Herald Sq - 34th St --> Jackson Hts - Roosevelt Av
     
   We can use above examples for nice representation.  